### 2021-01-17 FVT Manuscript/manuscript.Rmd
V5

---
title             : "Trait plasticity and covariance along a continuous soil moisture gradient"
shorttitle        : "Function-valued drought extremes"

author: 
  - name          : "J. Grey Monroe"
    affiliation   : "1,4"
    corresponding : yes    # Define only one corresponding author
    address       : "150 California Ave, Davis, CA 95616"
    email         : "gmonroe@ucdavis.edu"
  - name          : "Haoran Cai"
    affiliation   : "2"
  - name          : "David L. Des Marais"
    affiliation   : "2,3"
affiliation:
  - id            : "1"
    institution   : "Department of Plant Sciences, Univeƒrsity of California at Davis, Davis, USA"
  - id            : "2"
    institution   : "Department of Civil and Environmental Engineering, Massachusetts Institute of Technology, Cambridge, MA, USA"
  - id            : "3"
    institution   : "The Arnold Arboretum of Harvard University. Boston, MA, USA"
  - id            : "4"
    institution   : "Max Planck Institute for Developmental Biology, Tubingen, DE"

abstract: Water availality is perhaps the greatest environmental determinant of plant yield and fitness. Yet our understanding of plant-water relations is limited because it is primarily informed by experiments considering soil moisture variability at two discrete levels - wet and dry - rather than as a continually varying environmental gradient. Here we used experimental and statistical methods based on function-valued traits to explore responses to continuously varying soil moisture gradient in physiological and morphological traits in two species and five genotypes each of the model grass \textit{Brachypodium}. We find that most traits exhibit non-linear responses to soil moisture variability. We also observe differences in the shape of these non-linear responses between traits, species, and genotypes. Emergent phenomena arise from this variation including changes in trait correlations and evolutionary constraints as a function of soil moisture. These results point to the importance of considering non-linearity in plant-water relations to understand plastic and evolutionary responses to changing climates.

keywords          : "drought, function-valued traits, Brachypodium, evolutionary constraint, phenotypic plasticity, non-linearity"
wordcount         :

bibliography      : ["r-references.bib"]
csl               : mycsl.csl

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
figsintext        : yes

documentclass     : "apa6"
classoption       : "jou"
output            : papaja::apa6_pdf

header-includes   : 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r setup, include = T, warning=FALSE}
library("papaja")
library("ggplot2")
library("cowplot")
library("tidyverse")
```

```{r, message=F, results="hide", warning=FALSE}
source("../src/prepare_data.R")
source("../src/plot_fvt_byspecies.R")
source("../src/calc_fvt_byspecies.R")
source("../src/corplots_byspecies.R")
source("../src/constraints_byspecies.R")
source("../src/drydown.R")
source("../src/maps.R")
```

# Introduction

For plants, soil water availability is an important environmental factor in ecology and agriculture, acting as a major determinant of fitness and yield [@juenger2013natural; @greenham2017temporal]. As such, considerable effort has been placed on studying plant responses to drought, often defined conceptually and experimentally as an environmental condition of abnormally elevated aridity resulting in decreased plant performance [@passioura1996drought]. Most of the research on drought responses and tolerance strategies, including in _Brachypodium_, the focal system of this work, has involved comparisons between discrete soil water levels - control and water-limited [@des2012physiological; @edwards2012quantitative; @el2014genotype; @greenham2017temporal; @vasseur2014multivariate]. Yet soil moisture as an environmental factor is complex and multidimensional, with fluctuations varying continuously in timing, duration, and degree. Here we investigate trait responses to one important dimension of soil moisture variability - degree - with experimental and statistical approaches treating soil moisture content as a continuous variable rather than a set of fixed levels.

Quantitative genetic variation in drought resistance traits have been observed in natural populations and laboratory model systems. In particular, natural populations of Brassicaceae species (including _Arabidopsis thaliana_) and _Brachypodium_ harbor variation both constitutive and plastic traits mediating plant-water relations, including water use efficiency [@des2012physiological; @edwards2012quantitative; @greenham2017temporal; @des2017interactive], leaf chemistry [@des2017interactive; @kesari2012intron], leaf anatomy [@skirycz2011pause; @verelst2013molecular;@dittberner2018natural], root-shoot biomass partitioning [@des2012physiological; @des2017interactive], and many others [@verslues2011drought; @edwards2012quantitative; @juenger2013natural; @yarkhunova2016selection; @luo2016specific; @lenk2019transcriptional]. Among the many plant traits that can and have been measured, few have been studied more extensively than specific leaf area (SLA - often reported as its inverse, leaf mass per area or LMA). SLA provides a description of leaf architecture that is central to the leaf economics spectrum, a theory which seeks to explain variation in leaf physiological strategies, from more conservative (low SLA) to more productive (high SLA) [@wright2004worldwide]. In the context of drought stress, low SLA might be adaptive as lower leaf surface area is expected to reduce water loss through transpiration. In several cases, reductions in SLA have been reported under drought conditions [@casper2001drought]. However, for all of these traits, the shape of plastic responses to variation in water availability remain largely unmeasured.

Less is known about how trait variation and relationships between traits change across continuous environmental gradients [but see @robinson2009impact]. Genetic variation for traits can be higher, for example, in less frequently encountered environmental conditions [@gibson2004uncovering; @schlichting2008hidden; @paaby2014cryptic]. And describing this structure is important as trait variances and covariances influence evolutionary constraints. For example, if total genetic variance in trait space changes depending on the environment, then the capacity to respond to selection will vary accordingly, with reduced responses to selection under conditions where genetic variance is lower and vice versa. Similarly, if trait covariances depend on environment [@wood2015environmental], conditions which increase trait covariances may limit evolutionary potential across a range of environments by reducing the effective axes of variation [@levins1968evolution; @lande1983measurement;  @via1985genotype; @kingsolver2003environmental; @gomulkiewicz2018variation]. These phenomena are made more complex by the possibility that the relationship between trait variances and covariances with the environment may be non-linear. Investigating the genetic architecture of multiple traits is therefore useful for understanding evolution in rapidly changing environments. Fortunately, this need has been met with the development of statistical methods to reduce the dimensionality of genetic variance covariance matrices and produce meaningful summaries describing evolutionary constraints of multiple phenotypes simultaneously [@houle1992comparing; @blows2007tale; @kirkpatrick2009patterns; @kingsolver2015genetic].

To study the shape of phenotypic responses to the environment, modeling traits as mathematical functions of continuous variables, or function-valued-traits, has proven powerful [@kirkpatrick1989quantitative; @kingsolver2001variation; @griswold2008hypothesis; @stinchcombe2012genetics; @goolsby2015phylogenetic; @gomulkiewicz2018variation]. Recent decades have seen this approach used across many organisms [@gibert1998describing; @pettay2008age; @rocha2012connecting; @mason2020learning], traits [@robinson2009impact; @stinchcombe2010across; @mcguigan2010quantitative], and components of the environment [@brommer2008exploring; @mcguigan2009condition; @pearse2019life]. For plant scientists, experimental and analytical approaches leveraging the concept of function-valued-traits provide a framework for gaining a deeper understanding of plant acclimation and evolutionary adaptation to the environment.

In this study we combined these approaches into an exploratory investigation of trait plasticity, covariance and evolutionary constraints across soil moisture gradients. We model trait responses to a continuous gradient of soil moisture for multiple genotypes of two species of the model grass genus _Brachypodium_:  annual _B. distachyon_ and perennial _B. sylvaticum_. Because the shapes of trait responses cannot necessarily be known _a priori_, we use model selection among linear and non-linear environmental predictors to estimate the response function for each trait. We then estimate genotype means at different levels of soil moisture and compute the variance-covariance parameters for all traits. Finally, we ask whether patterns of variance and covariance in drying-responsive traits in _Brachypodium_ species may lead to variation in evolutionary constraint as a function of soil moisture.

# Materials and Methods

### Genotypes and species

_Brachypodium_ is a model genus for the genetics and genomics of C3 grasses [@brkljacic2011brachypodium]. In this work we studied natural variation between and among two species of _Brachypodium_: the annual _B. distachyon_ and the perennial _B. sylvaticum_. Both species are endemic to Eurasia, with _B. distachyon_ more prevalent in seasonally dry habitats in Southern Europe, North Africa, and the Middle East, and _B. sylvaticum_ more widely distributed throughout Eurasia [@catalan2016phylogeny] (Figure \@ref(fig:maps)). The different evolutionary histories, ecological environments, and life habits of these two species may have impacted patterns of genetic variation and evolvalbility in relation to soil moisture.

To investigate genetic variation, five genotypes of each were studied to characterize patterns of variation in plant traits across an environmental gradient: _Brachypodium distachyon_ inbred lines ABR2, Adi-10, Bd21, Bd3-1, and Koz-1 and _Brachypodium sylvaticum_ inbred lines Ain-1, Ast-1, Kry-1, Osl-1, Vel-1. For each species, these genotypes represent a range of geographical origins and phenotypic diversity (Personal observations) [@steinwand2013sylvaticum; @des2017interactive]. For example, Kry-1 is native to Krym, Ukraine while Ain-1 is native to Ain Draham, Tunisia [@steinwand2013sylvaticum]. Both species are self-compatible and each of the lines used here have been maintained as inbred lines for greater than six generations [@steinwand2013sylvaticum; @vogel2009development]; as such, experimental replicates may be considered nearly isogenic.

### Plant growth and dry down experiment

(ref:drydown) **Effect of the experimental dry down on soil water content.** (A) Time series of gravimetric soil moisture for all pots during the 14-day dry down period. (B) Distribution of final (day 14) soil moisture content across all pots. The data are distinguished by color according to the watering treatment. (C) Final soil moisture content by genotype.

```{r drydown, fig.cap='(ref:drydown)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/DryDown_lines.pdf", dpi = 108)
```

Plant growth and experimental soil dry down were performed in the greenhouses of the Arnold Arboretum of Harvard University. To synchronize germination across genotypes within each species, seeds were placed on damp filter paper in the dark at 4°C for 14 days prior to planting. To synchronize the developmental stage at the timing of the drought treatments between the two species, _B. sylvaticum_ seeds were planted thirteen days before _B. distachyon_ (Oct 7 and 20, 2015, respectively). For each genotype, 1200 seeds were planted two to a pot and were subsequently thinned to one plant, for a total of 600 experimental plants in a randomized block design. All plants germinated within four days of sowing. Individual seeds of plants were sown in Greens Grade Profile porous ceramic rooting media (Profile Products, Buffalo Grove, IL, USA) in Deepot D40H Conetainers (650mL; Stuewe & Sons. Tangent, OR, USA) and grown at 25°C/20°C days/nights. Ambient sunlight was supplemented to 1000 umol/m2/s for 12hr/day.

Dry down treatments began 29 and 42 days after sowing (DAS) for _B. distachyon_ and _B. sylvaticum_, respectively. Because the harvesting was divided over five consecutive days (see section below), plants were split into five equal harvest cohorts, with each cohort containing equal numbers of each watering treatment to avoid confounding harvest day with soil moisture content. Thus, though each consecutive cohort differed in age by a single day, each experienced the dry down treatment for the same amount of time. Nevertheless, we expected the difference in age between harvest cohorts to potentially impact trait expression and we therefore included harvest day (cohort) as a covariate in subsequent models. To generate a continuous gradient of final soil moisture by the end of the dry down period, plants were split into five watering treatments, receiving approximately 0, 4, 8, 12, 16, or 20 ml of water per day for 14 days (Figure \@ref(fig:racks)). Prior to initiating the experiment pots were weighed with dry soil ($mass_{dry}$) and field capacity ($mass_{max}$). During the course of the dry down, soil moisture content was calculated during the morning of day $d$ for each pot as $mass_d/(mass_{max}-mass_{dry})$. The final soil moisture reflects the combined effects of water input and output through evaporation from the soil and evapotranspiration from the plants.

### Plant harvesting and phenotyping

To characterize phenotypic responses to our experimental soil moisture gradient, we measured a suite of developmental and physiological traits. Plants were harvested in five cohorts over five days at the end of the dry down period. Each day, half of the sampled plants were harvested for above and below ground biomass, total above ground green area, $\delta C_{13}$, N content, C content. The other half were harvested and assessed for specific leaf area (SLA) and relative water content (RWC). Above ground leaf area was estimated by laying freshly harvested plants flat between plates of clear acrylic and imaging with a Nikon 5300 digital camera at a fixed distance with a 35mm Nikkor lens. Total green pixels were counted for each image with Easy Leaf Area [@easlon2014easy] (https://github.com/heaslon/Easy-Leaf-Area) with settings shown in Figure \@ref(fig:leafarea). Above ground biomass was measured after drying leaf material overnight at 60°C and then for several weeks at room temperature. Below ground biomass was measured after washing the soil matrix from roots and drying them overnight at 60°C and then for several weeks at room temperature. Above and below ground biomass was measured after leaves and roots were dried. Leaf tissues for $\delta C_{13}$, $\delta N_{15}$, nitrogen (hereafter "N") content, and carbon (hereafter "C") content were ground to a fine powder and processed by the UC Davis Stable Isotope Facility. SLA was calculated by scanning the two youngest fully emerged leaves with a 1 $cm^2$ red square. Leaf area in $mm^2$ was calculated from these same images using Easy Leaf Area. SLA was calculated as $leaf\ area / biomass_{dry}$. These leaves were also used to calculate RWC. Prior to drying, fresh leaves were weighed ($biomass_{fresh}$) and then submerged under water in 15mL falcon tubes for several hours. They were then weighed ($biomass_{turgid})$, oven-dried overnight, and weighed again ($biomass_{dry}$). RWC was calculated as $(biomass_{fresh}-biomass_{dry})/(biomass_{turgid}-biomass_{dry})$.

```{r models, results = 'asis', echo = F}

pretty_table <- capture.output(apa_table(
 t_model_selection_summaries_byspecies
 , caption = "Model selection. 'x' indicates parameters chosen in best minimum AIC model. H = harvest day, G = genotype, E = final soil moisture, E^2 = quadratic parameter, s(E) = 2nd degree natural spline parameter",
 digits=3,
 font_size="scriptsize",
 longtable = F,
 landscape=F,
 align="c",
 midrules=1:nrow(t_model_selection_summaries_byspecies)
))

pretty_table <- gsub("\\{table\\}", "\\{table\\*\\}", pretty_table)
cat(paste(pretty_table, collapse = "\n"))

```

## Analyses
We used R for all statistical analyses. Code and data to generate this manuscript can be found at https://github.com/greymonroe/brachypodium_fvt.

### Function-valued traits
For the purposes of modeling phenotypic responses to variation in soil moisture content, we considered soil moisture content as the final soil moisture on day 14 of the dry down period for each plant, referred to in figures as $Soil\ moisture (\%_{final})$. A major challenge in studying function-valued traits is model selection. That is, identifying the functions that best describe the curvature (or lack thereof) in the shape of phenotypic responses to environmental gradients. Quadratic and natural splines have been suggested as potential functions to model non-linearities [@meyer2005randomregression], but assuming the appropriate function is problematic. Akaike information criterion (AIC) selection based on contrasting multiple complex models offers an effective means to balance predictability with over-fitting [@griswold2008hypothesis; @bolstad2010direct; @gomulkiewicz2018variation]. Because we sampled only 5 genotypes per species [@van1995linear] we treat genotype as a fixed effect in our model, which differs from random regression used elsewhere where genotypes are treated as random effects [@nussey2007evolutionary]. Thus, we began with the full fixed-effects linear regression model for traits as described below.

$Trait = H+G+E+E^2+s(E)_{df=2}+G*E+G*E^2+G*s(E)_{df=2}$

where $H = harvest\ day, \ G = genotype,\ E = soil\ moisture,\ E^2 = quadratic\ parameter,\ s(E)_{df=2} = second\ degree\ natural\ spline\ parameter$

We then selected a model for each trait using stepwise AIC model selection with the $stepAIC$ function from the package $MASS$ [@MASS] in R with the "direction" parameter set to "both." The two species were analyzed separately to avoid biases introduced by enforcing the same model on species with different sizes, developmental trajectories and evolutionary histories.

### Genetic correlations as a function of soil moisture content
We calculated trait correlations at different levels of soil moisture to characterize how genetic correlations between traits vary as a function of soil moisture content. Predicted genotypic means for each trait were calculated at 20 levels of soil moisture content (from 30% to 100% gravimetric water content) based on the model chosen by AIC (see above). Pearson correlation coefficients between genotype means were calculated at each level of soil moisture within each species.

### Plasticity through multidimensional trait space
We quantified total plasticity through multidimensional trait space as a function of soil moisture by first scaling each trait to a mean of 0 and standard deviation of 1. We then calculated the euclidean distance matrix between genotype means at all soil moisture levels. We then estimated total plasticity, measured as euclidean distance between trait values, between consecutive soil moisture levels for each genotype. $m$ is a vector of length 20 evenly spaced soil moisture measurements from 30% to 100%:

$$m=(m_1,...,m_{20})=(30\%,...,100\%)$$

If $p_{g,i} = (p_{g,1},...,p_{g,n})$ and $p_{g,i+1} = (p_{g,i+1,1},...,p_{g,i+1,n})$ are two consecutive (in terms of soil moisture) points in $n$ dimensional trait space of scaled genotype means, total plasticity (Euclidean distance) for genotype $g$ across $n$ traits at soil moisture level $i$ is calculated as

$$\Delta T_{g,i}=\sqrt{\sum_{j=1}^{n}(p_{g,i,j}-p_{g,i+1,j})^2}$$

Finally, to ask if total plasticity differed between the two species, at each level of soil moisture, we then compared the mean $\Delta{T}$ between _B. sylvaticum_ and _B. distachyon_ by T-tests. To visualize plasticity of each genotype through multivariate trait space further, we performed a principal component analysis from the matrix of scaled genotype trait means using the _prcomp_ function in R.

### Analysis of evolutionary constraints among traits

We calculated several statistics summarizing evolutionary constraint as described in @kirkpatrick2009patterns. First, for each species, we approximated the $G$ matrix of genetic covariances by calculating variances and covariances between mean scaled genotype trait means at different levels of soil moisture. We then calculated, using the _prcomp_ function in R, the eigenvalues of each mean standardized (trait values divided by mean) $G$ matrix, $\lambda_i$. From these we then calculated the _number of effective dimensions_, $n_{D}$, equal to the sum of the eigenvalues divided by the largest eigenvalue:

$$n_{D} = \sum_{i=1}^{n} \lambda_i/\lambda_l$$

We also calculated the _maximum evolvability_, $e_{max}$, equal to the square root of the largest eigenvalue, $\lambda_l$ [@kirkpatrick2009patterns; @houle1992comparing]:

$$e_{max} = \sqrt{\lambda_l}$$

Finally, we calculated the _total genetic variance_ [@kirkpatrick2009patterns], equal to the sum of the eigenvalues of $G$:

$$v_T = \sum_{i=1}^{n} \lambda_i$$

(ref:curves) **Variation in phenotypic responses to soil moisture gradient as function-valued traits.** Points and the lines connecting them indicate predicted values at different levels of gravimetric soil moisture content from 30% to 100%. _B. sylvaticum_ genotypes are colored orange and _B. distachyon_ blue. (a) `r real_trait_names[1]` (b) `r real_trait_names[2]` (c) `r real_trait_names[3]` (d) `r real_trait_names[4]` (e) `r real_trait_names[5]` (f) `r real_trait_names[6]` (g) `r real_trait_names[7]` (h) `r real_trait_names[8]` (i) `r real_trait_names[9]` (j) `r real_trait_names[10]` (k) `r real_trait_names[11]` (l) `r real_trait_names[12]`

```{r curves, fig.cap='(ref:curves)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../Figures/curves_byspecies.pdf", dpi = 108)
```

# Results

### The dry down experiment resulted in a continuous soil moisture gradient

Across the six watering treatments, combined with random variation in water capacity of pots (Figure \@ref(fig:pots)), the dry down period resulted in a continuous environmental gradient of final soil moisture, albeit with a higher frequency of plants near the driest extreme of soil moisture variation (Figure \@ref(fig:drydown)). This gradient provides the basis for analyzing phenotypes in relation to soil moisture treated as a continuous gradient rather than limited set of discrete factors.

The observed reduction in leaf relative water content (RWC) under the driest conditions in both _B. distachyon_ and _B. sylvaticum_ indicates that at this extreme, plants were physiologically stressed (Figure \@ref(fig:curves)A). Mean leaf RWC for plants in the 10% tail of soil water content was `r tailrwc`% which is drier than that observed in the less severe soil drying conditions of Des Marais et al. [@des2017interactive]. Additional visual observations made during the experiment such as leaf rolling, another symptom of dehydration stress, were evident in plants at the lowest water treatment by the end of the dry down period.

(ref:multi) **Plasticity through multivariate trait space.** (a) Plasticity across all traits was calculated as distance between scaled phenotype for each genotype between different levels of soil moisture. Box plots indicate species median and 25th and 75th percentiles with whiskers extending to 1.5 times the interquartile range. "*" indicates significant differences between species (t-test, $\alpha < 0.05$). _B. sylvaticum_ is colored in orange and _B. distachyon_ blue.

```{r multi, fig.cap='(ref:multi)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../Figures/multi_plast.pdf", dpi = 108)
```

### Non-linearity in trait responses to soil moisture is common

We evaluated the degree to which traits showed linear or non-linear responses to soil moisture using an AIC model selection approach from a full model which included quadratic and natural spline parameters relating soil moisture content to plant phenotypes. We observed non-linear components (quadratic, spline, or both) in the final models for all of the traits which included an environmental (water content) predictor (Table \@ref(tab:models), Figure \@ref(fig:curves), Table \@ref(tab:outmodels), Table \@ref(tab:modelsr2)). In _B. distachyon_, all of the traits included at least one non-linear environmental predictor. In contrast total biomass and root mass did not include an environmental predictor in selected models in _B. sylvaticum_.

When considering plasticity across multidimensional trait space, it appears that most of the variation in _B. distachyon_ is attributable to responses to low soil moisture (Figure \@ref(fig:multi), Figure \@ref(fig:pca)). In contrast, _B. sylvaticum_ was more responsive to extreme wet conditions than _B. distachyon_. We observed, particularly in _B. sylvaticum_, that phenotypes were similar between extreme dry and extreme wet soil moisture contents (Figure \@ref(fig:pca)b). This similarity may be explained by the quadratic parameters of trait functions where the curvature of trait responses can lead to similar phenotypes at both environmental extremes.

(ref:cors) **Relationships between trait correlations and soil moisture content.** Each point is a correlation coefficient calculated across five genotypes between the intersection of the two traits labeled on the diagonal. Correlations were calculated among genotypes by species (blue = _B. distachyon_, orange = _B. sylvaticum_).  Correlations are not shown for traits in species where genotype was not included in final model (Table 1).
```{r cors, fig.cap='(ref:cors)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/corsplots_byspecies.pdf", dpi = 108)
```

### Most traits show significant genetic variation

We also tested whether there was significant natural variation for the traits measured between genotypes in each species by looking at the parameters in the final model for each trait. Interestingly, across species, most traits (21/24) included a genotype term in the final model, indicating significant differences between genotypes in magnitude of traits across all levels of soil moisture (Table \@ref(tab:models), Figure \@ref(fig:curves)). Though not formally tested, for other traits there are clear distinctions between the two species. For example, $\delta C_{13}$ was considerably higher in _B. sylvaticum_ (Figure \@ref(fig:curves)). For SLA, while _B. distachyon_ showed a strong response to soil moisture, especially under the driest conditions, SLA in _B. sylvaticum_ was much less responsive to soil moisture. In contrast, _B. sylvaticum_ appears to show a more dramatic response in leaf composition estimated by N content and C:N ratio. We note too that most of the traits included an effect of harvest cohort; this is expected due to difference in age (up to 5 days, 9-11% total) of different harvest cohorts.

### Several traits show interactions between genotype and non-linear responses to soil moisture

Significant interactions between genotype and environmental parameters in a final model indicate the presence of genetic variation for plasticity (GxE) [@via1985genotype]. For those GxE interactions where the environmental parameter is non-linear, significant GxE indicates genetic variation for the shape of reaction norms. SLA and root mass showed a significant interaction between genotype and soil moisture in _B. distachyon_. In _B. sylvaticum_, RWC, N content, leaf green area, C:N ratio, and root to shoot ratio all showed interactions between genotype and soil moisture (Table \@ref(tab:models), Figure \@ref(fig:curves)). In each of these cases, the interaction between genotype and environment involved a non-linear environmental predictor, indicating not only variation for the magnitude of plasticity (i.e. slope) but also variation in the shape of responses.

(ref:constraints) **Contrasting patterns of evolutionary constraint between _B. distachyon_ and _B. sylvaticum_.** Summary statistics of evolutionary constraint as a function of soil moisture in _B. distachyon_ (blue) and _B. sylvaticum_ (orange). (a and b) The number of effective dimensions, $n_D$, estimates number of unconstrained axes of variation. (c and d) The maximum evolvability, $e_{max}$, corresponds to the square root of the largest eigenvalue of the genetic covariance matrix. (e and f) The total genetic variance, $v_T$, is equal to the sum of the eigenvalues of the genetic covariance matrix.
```{r constraints, fig.cap='(ref:constraints)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../Figures/constraints_byspecies.pdf", dpi = 108)
```

### Correlations between traits change as a function of soil moisture, often in a non-linear fashion

We calculated correlations between genotype trait means across soil moisture for traits where genotypic differences were observed (i.e. genotype predictor in trait models). Certain traits were tightly correlated regardless of environment. For example, correlations near 1 were observed between biomass and green area in both species across soil moisture. More complex relationships between trait correlations and soil moisture are observed in other trait combinations. For traits with genotype by non-linear environment interactions (Table \@ref(tab:models)), trait correlations showed non-linear relationships with soil moisture as well. Because more of these interactions were found in _B. sylvaticum_ the number of trait combinations showing non-linear relations between correlations and soil moisture appears to be higher than in _B. distachyon_ (Figure \@ref(fig:cors)). The correlation between C:N ratio and root:shoot ratio in _B. sylvaticum_, for example, varied from approximately 0.3 under the wettest environment to approximately -0.7 under the driest environment.

### Evolutionary constraints differ as a function of soil moisture and show contrasting patterns between _Brachypodium_ species

To assess evidence of evolutionary constraint on the sampled traits, we approximated and analyzed parameters of the genetic covariance matrix, $G$, in each species across the soil moisture gradient. These analyses revealed contrasting patterns of evolutionary constraint both in relation to soil moisture and between _B. distachyon_ and _B. sylvaticum_. In _B. distachyon_, the number of effective dimensions ($n_D$, which estimates number of axes of variation unconstrained by covariance) was lower when soils were drier. In contrast, $n_D$ was lower in _B. sylvaticum_ when soils were wetter. The maximum evolvability ($e_{max}$, variance through largest eigenvector of multidimensional trait space) also showed opposite trends between the two species. Whereas in _B. distachyon_ $e_{max}$ was highest under the driest conditions, in _B. sylvaticum_ $e_{max}$ was highest under the wettest conditions. The same trend was seen in total genetic variance ($V_T$, which summarizes all genetic variance through multidimensional trait space).

# Discussion

### Non-linearity in soil moisture response is common in _Brachypodium_

We found significant non-linearity in response to a soil moisture gradient for all measured traits in at least one of the two species sampled. The best-fit function for some traits was quadratic, while other traits showed more complex responses to the environment which were best fit by a spline function. These results offer new insights with respect to the study of plant response to soil drying. By focusing on the curvature of phenotypic response as the explicitly modeled trait, we avoid contrasts of trait values expressed at arbitrary levels of soil water content which may obscure different thresholds of response among the diverse genotypes under study. SLA in _B. distachyon_ exhibits this pattern, as two accessions show a threshold-like response in decreasing SLA as soils become drier, and three accessions express their highest SLA at intermediate soil water content. Leaf N content (on a leaf-mass basis) in _B. sylvaticum_ likewise shows considerable diversity of response with two accessions expressing their lowest values at intermediate soil water content, one accession expressing its highest leaf N at intermediate soil water content, and one accession showing a nearly linear response along the soil water content gradient.

Non-linearities in trait responses to soil moisture reinforce the need to consider the consequences of extreme weather events when predicting plastic responses, especially when scaling up to investigating the ecosystem effects of individual plant responses to environmental stress [@felton2017integrating]. As others have previously noted, if organism responses to the environment show non-linear curves then changes in the environment may result in greater or lesser responses than expected based on predictions based on linear response curves alone [@nussey2007evolutionary; @brommer2008exploring; @visser2008keeping; @gienapp2014evolutionary]. This suggests, for example, that responses of ecosystem traits such as SLA to increasing aridity and drought may depend on the severity of environmental change in a threshold-like fashion. Moreover, because traits like SLA have impacts on broader processes including carbon cycling,  carbon models incorporating plastic and evolutionary responses may benefit from information about the (non)-linearity of trait responses to environmental change [@donovan2014ecological; @monroe2018ecoevolutionary; @walker2019plastic].

### Implications for evolution of _Brachypodium_

Leaf N and SLA are two axes of the classic Leaf Economic Spectrum [@wright2004worldwide] and so the contrasting responses of these traits between the annual _B. distachyon_ and perennial _B. sylvaticum_ may reflect broader differences in their life history strategies. We recently reviewed evidence for physiological, anatomical and developmental differences between herbaceous annual and perennial species, finding support for generally higher SLA in annuals, befitting a generally resource-acquisitive strategy [@lundgren2020lifehistory]. Garnier [@garnier1992growth] argued that small changes in leaf anatomy (e.g. SLA) will likely have large effects on plant growth rate and resource use and could therefore tip the balance between perennial and annual strategies.

We also found that signatures of evolutionary constraint differ along our imposed soil water content gradient. Specifically, we find evidence of highest evolvability in multitrait space (measured by $e_{max}$ and $v_T$) in _B. distachyon_ under the driest soils. In contrast, _B. sylvaticum_ exhibited evidence of greater evolvability by the same measures under the highest soil water content studied here. However, _B. distachyon_ exhibits evidence of higher constraint caused by genetic covariances (low $n_D$ under dry conditions) These results indicate that _B. distachyon_ has increased genetic variance under dry conditions as compared to wet conditions, but that natural selection may be more constrained to act on this variation due to covariance between traits. In contrast, our results suggest that _B. sylvaticum_ has decreased genetic variance under dry conditions on which selection might act but that this variation is less constrained by covariance between traits.

Consistent with predictions of elevated genetic variation revealed in less frequently encountered environments [@gibson2004uncovering; @schlichting2008hidden; @paaby2014cryptic], we speculate that the pattern observed could be a reflection of the different life history strategies of these two species. Annuality is considered a drought adaptive strategy characterized by escaping drought through phenology, by flowering before and remaining dormant as seeds during the most drought prone seasons [@friedman2015all; @monroe2019drought; @friedman2020evolution]. Thus, because of their life history, populations of annuals may actually experience fewer episodes of strong selection from extreme drought, which could explain why we find elevated genetic variance under these environments. In contrast, perennials such as _B. sylvaticum_ are subjected to all seasons and might therefore experience more frequent episodes of selection caused by dehydration stress, despite paradoxically being in found in environments where droughts are less frequent on an annual basis. This pattern is consistent with the predictions of cryptic genetic variation revealed under environments where selection is less frequent or severe [@gibson2004uncovering; @schlichting2008hidden; @paaby2014cryptic]. We remain cautious about drawing this conclusion from our data, as they represent a comparison between only two species and include a small sample genotypic diversity in each. However, that the difference in patterns of constraint are explained by alternative life history strategies emerges as an interesting hypothesis from these observations, one that might be addressed by future work extending the approaches here to a broader phylogenetic comparison between annual and perennial species.

### Potential to scale up investigations of function-valued responses to soil moisture

We found that genetic correlations between traits can vary over relatively small changes in soil moisture (Fig. \@ref(fig:cors)). Similarly, we found that signatures of evolutionary constraint varied across the environmental gradient, suggesting that responses to selection may be improved or restricted in accordance with patterns of constraint in relation to environment. Interestingly, we also found that some species may be more responsive to selection in a given environment based on patterns of constraint. However, in this study we focus on only five genotypes per species. This reflects the inherent challenge of investigating large numbers of genotypes and high resolution environment gradients. But larger-scale investigations are needed if results are to be extended to the large populations typical of genetic mapping studies and breeding pools. Indeed, responses to selection for drought tolerance may depend on drought severity because of these patterns in genetic correlations. And in the context of breeding, larger studies applying these approaches might be valuable for identifying conditions for which genetic correlations are aligned with breeding objectives [@bernardo2020breeding].

From a practical perspective this work highlights the potential of function-valued trait approaches that may be scaled up to studying plant-water relations. In this experiment, we investigated variation in plant responses to a gradient of soil moisture using six watering levels, which in combination with random variation in water capacity of experimental pots, produced a continuous gradient of soil moisture ranging from field capacity of the soil to strongly water-limited. In the field, multiple watering regimes in combination with random variability between plots may produce similar gradients of soil moisture. Here, water content was measured gravimetrically. New sensing technologies may be useful for quantifying soil moisture or other continuous environmental gradients in an analogous fashion, to define soil moisture quantitatively and then apply function-valued statistical approaches to contrast trait expression among genotypes. Finally, while in this experiment we used destructive phenotyping methods to measure traits, non-destructive (and high throughput) phenotyping enable measurement of yield or fitness data as well to examine explicit connections between trait variation and adaptation to different degrees of soil moisture in larger experiments. Together these approaches provide  opportunities to scale up the analytical framework used here to study genetic variation for soil moisture function-valued-traits in populations of diverse genotypes.

# Acknowledgements
We thank Chase Mason, Eric Goolsby, and Kyle Hernandez for insightful conversations about function-valued trait approaches. We thank Benoit Pujol and two anonymous reviewers for feedback on earlier versions of this manuscript. This work was supported by an Eco-Evo-Devo RCN training grant, USDA-NIFA Award 2014-38420-21801, and Max Planck Society support for JGM. Version 5 of this preprint has been peer-reviewed and recommended by Peer Community In Evolutionary Biology (https://doi.org/10.24072/pci.evolbiol.100119).

# Contributions
JGM and DLD funded, planned, and conducted the experiment. JGM, HC, and DLD contributed to analyses and writing.

# Conflict of interest disclosure
The authors of this preprint declare that they have no financial conflict of interest with the content of this article. JGM is one of the PCI Evol Biol recommenders.

\newpage

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

\newpage
\beginsupplement

# Supplement

```{r outmodels, results = 'asis', echo = F}

pretty_table <- capture.output(apa_table(
 out_models
 , caption = "final models",
 digits=5,
 font_size="tiny",
 longtable = F,
 landscape=F
)
)

pretty_table <- gsub("\\{table\\}", "\\{table\\*\\}", pretty_table)
cat(paste(pretty_table, collapse = "\n"))

```

```{r modelsr2, results = 'asis', echo = F}

pretty_table <- capture.output(apa_table(
 r2
 , caption = "Selected model r-squared and AIC values",
 digits=3,
 font_size="tiny",
 longtable = F,
 landscape=F
)
)

pretty_table <- gsub("\\{table\\}", "\\{table\\*\\}", pretty_table)
cat(paste(pretty_table, collapse = "\n"))

```

(ref:maps) Distributions of (a) _B. distachyon_ and (b) _B. sylvaticum_ reported on GBIF as of 2019.18.08. Examples of (c) _B. distachyon_ Carly Slawson (CC BY 4.0, https://www.inaturalist.org/photos/42532397) and (d) _B. sylvaticum_ Grzegorz Grzejszczak (CC BY-NC 4.0, https://www.inaturalist.org/photos/36088991) (GBIF.org (26 February 2020) GBIF Occurrence Download https://doi.org/10.15468/dl.rau5v9).

```{r maps, fig.cap='(ref:maps)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/maps.pdf", dpi = 108)
```

(ref:racks) Experimental design. Shown is the schema for each of the 10 genotypes. 
```{r racks, fig.cap='(ref:racks)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/design.png", dpi = 108)
```

(ref:leafarea) Settings used in Easy Leaf Area. 
```{r leafarea, fig.cap='(ref:leafarea)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/settings_used.pdf", dpi = 108)
```

(ref:pots) Variation in pot field capacity. 
```{r pots, fig.cap='(ref:pots)', out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("../Figures/pots.pdf", dpi = 108)
```

(ref:pca) **Plasticity through multivariate trait space.** Principal component analysis of scaled phenotypic responses to soil moisture gradient among genotypes of both species. Upper panels show genotype means across soil moisture content. Percent values in axis titles indicate percent variance explained by that principal component. Lower panels show eigenvectors of each trait. (a) PC1 and PC2. (b) PC3 and PC4. _B. sylvaticum_ is colored in orange and _B. distachyon_ blue.

```{r pca, fig.cap='(ref:pca)', out.width = "\\textwidth", fig.pos = "!h", fig.env="figure*"}
knitr::include_graphics("../Figures/pca_byspecies.pdf", dpi = 108)
```